Horizontal and vertical diversity drive total biomass and its temporal stability in fish communities

### 2020-05-26 manuscript/bef_stability/manuscript.Rmd
Fix PCA plot

---
title: Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems 
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

@@

@@

#### Biodiversity and ecosystem functioning

Biodiversity is essential to determine and maintain ecosytem properties.
Current biodiversity losses question the ability of how ecosystems will be
affected in turn.

The relationship between biodiversity and ecosystem functioning has been a
central research topic in ecology [@may_will_1972; @hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @duffy_why_2009;
@isbell_forest_quantifying_2018].

Studies in
Bioversity-Ecosystem
Functioning (BEF) usually quantify species diversity as species richness (i.e
number of species) while ecosystem functioning is quantified as total biomass,
productivity (biomass produced by unit of time) or stability which is quantified
in more diverse ways such as the variability of species abundance, the
resilience to perturbation, or resistance to perturbation
[@donohue_navigating_2016].

#### Shape of BEF: total biomass

Species richness has been reported to increase productivity
[@tilman_productivity_1996; @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partitioning_2001; @cardinale_species_2002; @fargione_selection_2007].
Empirical assessment of the effect of species richness on total biomass showed
negative [@grace_integrative_2016] and positive relationships
[@hector_plant_1999; @tilman_biodiversity_2014]. @grace_integrative_2016 showed
that the results depend on the refinement of the mecanisms involved in the
analysis.

#### Shape of BEF: stability

The relationship between species richness and stability is even more complex
partly because of the inherent multidimensionality of stability
[@donohue_dimensionality_2013; @donohue_navigating_2016] and because of the
different metrics of stability privileged in theoretical (stability of local
equilibrium, robustness to primary extinction) and empirical work (variability
of abundance or biomass through time). Empirical studies showed consistently
that species richness decreased community variability in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018] and
vertebrates [@olivier_independent_nodate; @franssen_annual_2011]. Two mecanisms
are involved into community variability, (1) species variability and (2) the
species synchrony [@loreau_species_2008; @thibaut_understanding_2013].  Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. On the other side, studies consistently found
that species richness decreased synchrony between species [@loreau_species_2008;
@thibaut_understanding_2013; @bluthgen_land_2016; @olivier_independent_nodate].
Beyond species richness, species interactions, which modifies species abondance
and biomass and make species dynamics interdependant, have been largely ignored
in empirical assessment.

#### BEF: Theoretical ecology

Conversely, theoretical ecology display an extensive literature about the
effects of network structure on stability [@may_will_1972;
@angelis_stability_1975; @stouffer_compartmentalization_2011;
@thebault_stability_2010; @barbier_pyramids_2019;
@shanafelt_david_w._stability_nodate].  Theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small perturbations.  Connectance (i.e. proportion of realized links in a
network) can decrease [@angelis_stability_1975; @thebault_stability_2010] or
increase resilience of trophic network depending on the functioning of networks
[@angelis_stability_1975].  The number of trophic levels, the vertical position
in the trophic network as well as the strength of their biomass regulation (i.e.
top-down and bottom-up) can determine biomass and stability at equilibrium
[@hurd_stability_1971; @shanafelt_david_w._stability_nodate;
@zhao_horizontal_2019; @barbier_pyramids_2019].

A large inconsistency remains between theoretical and empirical assessments of
the relationship between community structure and ecosystem properties.
Theoretical studies have ignored temporal variability and then the relative
contribution of synchrony and species variability on community variability while
empirical studies about temporal variability and total biomass in natural
communities have ignored the contribution of network structure.

Our goal was to assess the mecanisms determining temporal variability and total
biomass in natural communities while integrating both species richness and
network structure. We combined together the effect of vertical and horizontal
diversity and thus aimed to bridge the current gap between theoretical and
empirical studies. We used an standardized monitoring of fish communities in
french streams. More than three hundreds sites were followed more than 10 years
on the 1995-2018 period. We inferred four thousands potential interaction
networks corresponding to each sampling event, based on trait matching.

We investigated the effects of the community structure on total biomass and
community variability, which in turn determine community variability. We have
also included the effect of environmental drivers [@grace_integrative_2016]
which can directly affect community variability or indirectly via community
structure.

- Q1: diversity-stability, diversity-biomass in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3:

@@

## Community variability

@@

Fish communities were monitored across streams of France on the period
1995-2018 by a French governmental agency (OFB, previously AFB and ONEMA). We
kept only standardized protocols: complete and partial methods for respectively
little and big streams (detailed in Supplementary methods). In the complete
protocol, the sampling takes place over the whole width of the river and is done
by foot whereas in the partial protocol, the sampling is done over the bank only
and by boat. The sampling took place either in late spring and/or Autumn.
Heterogeneity in the sampling time may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, we select only one sampling per year if there
was two in a given year. Most of the fish sampling occurred in September
(Median: mid-August) and within station the median standard deviation of the
sampling month was inferior to one (0.7). Additionally, we removed sampling
events, in a given station, which the length of the stream sampled was outside
the median length sampled more or less 30%. The number of sampling events is
critical in order to estimate a meaningful variability of the community. We kept
only station that have been followed ten years or more (i.e 10 sampling events
or more), which sums to 403 stations.

The sampling itself was performed by electric fishing. All the stunned fish
species were identified and the individuals counted.

##### Biomass

To characterize communities variability, abundance or biomass are generally
used. When the body size and therefore body mass vary by several orders of
magnitude, one downside of abundance is that we mix individual that have
potentially really different meaning in terms of ecosystem properties such as
biomass. We therefore estimated the body mass of fish individuals according to
their body size. During sampling, the fishes were classified in batch according
to species and size class. Depending on the batches, the body size of all
individuals were measured or only a subsample. Only the body size big fishes was
measured individually. The body size of individual fishes in subsampled batches
was inferred under normality distribution assumption (detailed in Supplementary
material). The body mass of individual fishes were then estimated with an
allometric law ($B_i = 0.01 \times l_i^3.03$, $B$ being the body mass in grams,
$l$ the body size in millimeters and $i$ the individual).

The species and community biomass were then computed for each sampling event.
As sampling effort varied across protocol type and sampling event, biomass
was reported to the surface sampled (in $m^{2}$). Community and species biomass
were then expressed in $gm^{-2}$.

Total biomass was characterized as the median total biomass over time. Community
stability was defined as the inverse of biomass variability. Variability was
computed as Coefficient of Variation (CV) of the biomass ($CV_i = \sigma_i \mu_i^{-1}$,
with $\sigma_i$ and $\mu_i$ being respectively the standard deviation and the
average of the biomass of the station $i$). Then, biomass stability was equal to
$CV^{-1} = \mu \sigma^{-1}$.

The CV of community biomass can represent very different dynamics such as
decreasing, stable or increasing. However, the concept of community variability
refered to variability around an equilibrium, which is rarely observed or
difficult to characterize in natural communities. 
We However used SAX (Symbolic Aggregate Approximation) method to caracterize
rougly biomass dynamics over time (details in Supplementary information). We
finally kept only the stations which the biomass had no trends, remaining 99
stations.

To characterize the mecanisms driving biomass stability, we decomposed
the community CV in synchrony and weighted average species CV
components according to @thibaut_understanding_2013:

@@

with $CV_{com}$ being the Coefficient of Variation of the community,
$\overline{CV_{sp}}$ the average coefficient of variation of population
pondered by biomass. $\phi$ is the synchrony, which is the ratio the total
community variance ($\sigma^2_{x_T}$) and the sum of population variance
($\sigma_{x_i}$) [@loreau_species_2008]:

@@

@@

## Network building

@@

To obtain the network structure of the communities, we needed to infer
ecological interactions between the components of the communities and thus build
the ecological network. As the interactions were not measured, we used trait
matching method to infer the presence or absence of an interaction. This method is
widely used to infer trophic interactions
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. Indeed;
the existence of an trophic interaction between two species have been reported
to be largely dependant on their body size ratio
[cascade model; @cohen_stochastic_1985; @warren_invertebrate_1987; @brose_predator_2019].

Additionally to body size, we used food diet data to infer trophic
interactions. The diet of stream fishes, as many organisms, shifts over lifespan.
For example, alvins feeds on plankton while adults may feed on algae or fishes.
As we got body size of each fish individuals (cf Biomass inference methods and
supplementary material), we are able to define trophic interactions beyond
species level. We thus divided each fish species into nine body size classes,
each class having the same body size range (percentile method). Each body size
class constituting a trophic species. We then infered trophic interactions
between the trophic species. At this stage, it is important to make clear that
we infered potential trophic interactions and not actual measured interactions.
We determined network structure in two steps: (1) build a metaweb decribing the
trophic interactions between all the possible trophic species, then (2)
determine network of a given community by subtracting the metaweb according to
the trophic species present in a given community.

### Metaweb

#### Ontogenic diet

Diet information were obtained from database and literature [REF edeline?]. The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

#### Define trophic interactions

The metaweb described the potential trophic interactions between the $`r 57*9`$
($57\times9$) trophic species. We had seven resource nodes which were present
in the diet of the trophic species namely detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos.

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012].
For a given trophic species, we first determined if its size class end up in a
life stage containing piscivorous diet, i.e. if there was an overlap between
them. We were more liberal for piscivory determination to not miss potential
fish-fish trophic interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size range
from 3% to 45% of the median of its size range. A trophic interaction was set
for each trophic species whose median of its size class fell into the prey range
of the predator trophic species. To sum up, the metaweb was obtained with
allometric feeding interactions (1) between trophic species, (2) between trophic
species and resource nodes and (3) between resource nodes.

#### Obtain network of a community

The trophic network of a community (i.e. one network for each fish community
sampling) was obtained by determining the trophic species of each fish
individual in the community and interactions were set according to the metaweb.
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species. The inferred interactions represent potential interactions
(i.e. not measured interactions) and are binary (i.e. presence or absence of
feeding interactions).

#### Network structure

In each network, we computed two network metrics that have been commonly used to
characterize network structure. Connectance which describes the number of
interactions compared to the number of possible links,
 computed as ($C = L / N^2$, L and N being resp. the number of
links and of nodes). We then computed the average trophic level, the
average trophic level was weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass) ($\overline{T} =
\sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$ and $B_i$ being resp. the
trophic level and the biomass of the node $i$).

To sum up, we infered binary trophic networks at the scale of trophic species
(i.e. body size classes by species) to take in account ontogenic diet shifts and
that fish-fish trophic interaction depends on body size ratio. On the other
side, the community metrics were defined at the species level, (i.e. species
richness, average species CV and synchrony).

@@

## Habitat and environment

@@

Habitat description were obtained from fish sampling data. We kept variables
describing the ecosystem size namely the stream width, depth and the ecosystem
position (altitude and slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature as physical variables to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018; @tonkin_flow_2018; @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water.
The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects and took the annual mean (i.e. one
value by year and by station. As the environmental recording stations were not
located at the same place than the fish stations, we interpolated environmental
data with spatial statistical network model [@isaak_applications_2014]
implemented in the `SSN` R package [@hoef_ssn:_2014]. The interpolation
procedure is detailed in supplementary material. For each fish station,
environmental variables were matched with the year of fish sampling (except for
water temperature for which the record began in 2006). Then, the average
temperature, flow, BOD, river width and depth were computed as well as
coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and environmental
variables are often colinear, a principal component analysis (PCA) was performed
on the scaled variables. The PCA explained `r tot_var_expl_pca`% of the
variance. An axis rotation was performed to maximize the representation of the
variables by the principal components (Fig. \@ref(fig:pca-rotated)). We kept
five axis, representing respectively the (1) river size (river width & depth),
(2) the altitude, slope and decreasing average temperature, (3) average
enrichment (BOD) and CV of flow, (4) river size CV (CV of river width and depth)
and (5) CV of enrichment (CV of BOD). Additionally, we logged the axis one, two
and three to normalize their distribution and we reversed the third axis to get
increasing average temperature with increasing values of the axis.

@@

@@

### SEM

>>>>

---
title: Horizontal and vertical diversity drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, results = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

@@

```{r}
knitr::opts_chunk$set(
  results = TRUE 
)
```

@@

#### Biodiversity and ecosystem properties

Biodiversity is essential to determine and maintain ecosystem properties
[@duffy_why_2009].  Current biodiversity losses and global changes question the
way in which ecosystems will be affected in turn [@bellard_impacts_2012]. Since
1970s and the @may_will_1972 seminal paper, ecology aims to uncover the complex
relationships between biodiversity and ecosystem properties [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018]. Because biodiversity largely
determines both quantity and distribution of biomass or energy across
ecosystems, it has a tremendous effects on the properties exhibited by an
ecosystem, such as its total biomass [@hector_plant_1999;
@tilman_biodiversity_2014; @grace_integrative_2016;
@pennekamp_biodiversity_2018]], its biomass production
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016]
and the temporal stability of its biomass [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006;
@olivier_independent_nodate].

#### Horizontal and vertical diversity are two ways to look at biodiversity

Biodiversity can be quantified in numerous ways, as the number of species, i.e.
species richness and in number of functional groups, i.e species are
gathered on what they do. Those metrics can be gathered into "horizontal
diversity", describing the list of components present in a community, i.e. here
species and functional groups. An
orthogonal view, based on trophic chains, described species on what they eat. It 
can be defined as interactions (or biomass/energy fluxes) between species and how
they are distributed among communities. It can be quantified as the number
of interaction between species or the number of trophic level. This orthogonal
view of seeing biodiversity, can be gathered into "vertical diversity".  
Horizontal and vertical diversity are two views of quantifying biodiversity that
are, however, not independent from each other [@duffy_functional_2007; @zhao_horizontal_2019]. The add
or removal of one species can have different effects depending on its vertical
position in the trophic chain [@duffy_functional_2007]. Hence, it is crucial to
study their interplay to improve our understanding of ecosystem dynamics. 
(*This paragraph is surely too simplistic*)

#### horizontal and vertical diversity increase total biomass by common and different mechanisms

Horizontal and vertical diversity has been shown to promote higher biomass
production in a variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase productivity through complementarity between species in
resource exploitation and selection effects in plant communities
[@tilman_productivity_1996; @cardinale_species_2002; @grace_integrative_2016].
Regarding total biomass, empirical assessment have
shown generally positive relationships with species richness in plant
communities [@hector_plant_1999; @tilman_biodiversity_2014, but see
@grace_integrative_2016], as in mesocosms [@pennekamp_biodiversity_2018].
Theoretical models have also confirmed that an increase in horizontal diversity should
translate into increase in plant biomass through resource complementarity
[@loreau_biodiversity_1998] and selection effects [@tilman_plant_1997]. The
addition of trophic levels, notably consumer on plants generally breaks the
positive relationship between plant biomass and plant diversity as a significant
amount of biomass is transfered to consumers [@thebault_food-web_2003]. It has been shown both in
experiments [REF] and theoretical models [@thebault_food-web_2003].
@thebault_relationship_2006 have shown with a theoretical model that total biomass should increase with
the increase species diversity at two trophic level. However, the joint effects
of horizontal and vertical diversity on total biomass has been
rarely tested, especially in natural communities. It is not clear whether
horizontal and vertical diversity just change the distribution of biomass across
trophic level or it increases the total amount of  biomass (*It misses a lot of
references here*).

#### Horizontal and vertical diversity versus temporal stability

Horizontal and vertical diversity also affects the stability of ecosystem
properties. Experimental studies showed consistently that species richness
increase temporal stability of biomass in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018]. While in
natural communities, same relation has been found for temporal biomass stability
in stream fish communities [@franssen_annual_2011] and temporal abundance
stability of birds and butterflies [@olivier_independent_nodate, but not bats].
Two mechanisms are involved into temporal stability of communities, (1)
synchrony and (2) species variability [@loreau_species_2008;
@thibaut_understanding_2013]. Synchrony measures the temporal covariation of
species. Very low synchrony indicates compensatory temporal variation between
species and so increased temporal stability. Species richness has been generally
found to decrease synchrony between species [@loreau_species_2008;
@thibaut_understanding_2013; @bluthgen_land_2016; @olivier_independent_nodate,
but @declerck_species_2006]. The average temporal variability of species is
measured as a average coefficient of variation of species, weighted by species
abundance/biomass. Higher species temporal variabiliy thus results in lower
temporal stability. Positive [@tilman_biodiversity_2006], neutral
[@jiang_different_2009; @olivier_independent_nodate] and negative
[@franssen_annual_2011; @olivier_independent_nodate] relationships between
species variability and species richness have been found. This different
patterns can be dependent on species mean variance relationships and on the
relationship between evenness and species richness
[@thibaut_understanding_2013].  Synchrony has been found to be the main driver
of temporal stability [@tilman_biodiversity_2006; @olivier_independent_nodate]
and responsible for the positive relationship between horizontal diversity and
temporal stability of biomass and abundance. (*Should I add more details about
the mecanisms?*)

Vertical diversity, represented by network structure, has been central to the
literature of diversity-stability debate and largely explored by theoretical
ecology [@may_will_1972; @angelis_stability_1975;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. Because species
interactions make species dynamics interdependant, they have a tremendous role
on stability notably through its effects on the propagation of perturbations.
However, it is important to note that theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small independent perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the number
of secondary extinctions after species removal [@duffy_why_2009]. Those
stability metrics hardly relates to empirical temporal stability [@arnoldi_resilience_2016].
However, @thebault_trophic_2005 shows for a bi-trophic community
that asynchrony consistently increase temporal stability and a positive
relationship between temporal stability and species richness. However asynchrony
can also increase species variability through interspecific competition, which
amplifies species biomass fluctuations [@thebault_trophic_2005]. Furthermore
species interactions, considered as competition strength and niche breath of
consumers can results in negative or positive temporal stability-species
richness relationships
[@thebault_trophic_2005] depending on the food web configuration. Vertical
diversity, as network structure, can have strong effects on temporal stability.
However, there are few assesments of the effects of horizontal and diversity on
temporal stability in natural communities.

**Justify that horizontal and vertical diversity are linked**
Because horizontal & vertical diversity are not indedepent, they should be
analyzed jointly. For example, connectance generally covaries negatively with
species richness [@dunne_network_2006; @thebault_stability_2010; but see
@winemiller_must_1989]. This is also true for other metrics [@dunne_network_2006;
@delmas_analyzing_2018]. **TO DEVELOP, MOVE or RM**

#### Stream fish communities is a good model to study effects of horizontal & vertical diversity

Stream fish community constitutes a good model to study the joint effects of
horizontal & vertical diversity on total biomass and its temporal stability.
Despite that fish trophic interactions are difficult to assess in nature, a consequent
body of literature supports that they can be inferred through body size and diet
informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Most of the biomass-diversity and stability-diversity
relationships have been assessed in terrestrial ecosystems while there is a
evident need to test such relationships in a variety of ecosystem types. Stream
ecosystems are characterized by the flowing water and depends for a large part
from the input of terrestrial ecosystems. Furthermore, the biotic activity in
stream ecosystems determines for a large part the water quality
[@cummins_structure_1974] and thus are crucial to maintain **human selfish
goals (can do better than that)**.

#### Environment plays a big role in structuring communities

Environment also largely structure communities and is important to take in
account when considering natural communities. Environment directly affects
organism performance, and subsequently, abundance of species and in turn
determines total biomass [Add general ref; @grace_integrative_2016] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @friedpetersen_drivers_2020]. Hydrology of streams
constraints the composition of fish communities [REF toulouse à balle], and is
also related to fish population temporal stability [@franssen_annual_2011].
However it is not clear which is the relative strength of direct and indirect
effects of environment on total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

#### Research questions

The joint effects of horizontal & vertical diversity on total biomass and its
temporal stability remains yet poorly documented, especially in natural
communities. In this study, we questionned (1) whether the relationships between
horizontal diversity and total biomass/temporal stability are positive or
negative in aquatic ecosystems (Figure \@ref(fig:matmet) A). Then, as indicated
by theoretical studies, (2) what are the effects of vertical diversity on total
biomass and its temporal stability? Considering both direct & indirect effects,
(3) What is the magnitude of environmental effects on total biomass and its
temporal stability compare to horizontal and vertical diversity? (4) Which are
the pathways of the effects of environment, horizontal & vertical diversity on total
biomass and its temporal stability ? (4a) For environment, is it mainly (i)
direct, indirect (ii) through horizontal diversity, (iii) through horizontal &
vertical diversity or (iv) indirect through only vertical diversity? (4b) For
horizontal diversity, is it mainly (i) direct, or (ii) indirect through vertical
diversity?  Finally for temporal stability, (4c) how do environment, horizontal
& vertical diversity affect synchrony & species stability ?

#### SEM is a convenient approach to investigate direct and indirect path

To reveal the direct and indirect effects of environment, horizontal & vertical
diversity on both total biomass and its temporal stability, we built and
analyzed Structural Equation Models (SEM). It is one promising way to go beyond
bivariate relationships and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. SEM makes possible to set
causal relationships between variables. This step can be critical, as the
estimated parameters can be determined by both the choice of the variables and
the links between them. Therefore, the set of variables and links should be
supported by theory and literature. We built a metamodel which present the set
of variables and the links between them (Figure \@ref(fig:matmet) A,
\@ref(fig:sem-plot) A & B). Basically, we tested the links between environmental
conditions, horizontal & vertical diversity and both total biomass and its
temporal stability (Figure \@ref(fig:matmet) A). We considered that environment
can directly affect horizontal and vertical diversity, as well as total biomass
and its temporal stability. Horizontal diversity, measured as species richness
directly affects vertical diversity, total biomass and its temporal stability.
Food web structure was shown to be constrained by species richness. In turn,
vertical diversity affects total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

We used an standardized monitoring of fish communities in french streams. 99
sites were followed more than 10 years on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks, based on body size and food diet.
Horizontal diversity was quantified as species richness while vertical diversity
was as connectance and average trophic level. Environment was characterized by
the stream size, altitude, slope, and the temperature, flow, enrichment and
their temporal variability (expect for altitude and slope). Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

- Q1:  horizontal diversity-biomass horizontal diversity-stability in aquatic systems
- Q2:  in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3: Relative contribution of environment and community structure to biomass
  and stability

@@

@@

Fish communities were monitored across streams of France on the period 1995-2018
by a French governmental agency (Office Français de la Biodiversité), using
electric fishing. Two different standardized protocols were used for little and
big streams (detailed in Supplementary methods). For streams shallower than (XX
m), the sampling takes place over the whole width of the streams and is done by
foot whereas in streams deeper than (XX m), the sampling is done over the bank
only and by foot or by boat. The sampling took place either in late spring and/or Autumn
and temperal series were checked to reduce sampling heterogeneity within
stations (detailed in Supplementary methods). Each station have been followed
ten years or more (i.e 10 sampling events or more). The temporal stability of
biomass, computed as inverse of the CV, can cover very different dynamics such
as decreasing, stable or increasing. However, the concept of community
variability referred to variability around an equilibrium, which is rarely
observed or difficult to characterize in natural communities. We finally kept
only the 99 stations (Figure \@ref(fig:matmet) B) which had no trends in
biomass (method detailed in Supplementary methods). However, the results
were robust to the relaxation of this hypothesis (Supplementary results).

##### Total biomass and its temporal stability

To characterize communities variability, abundance or biomass are generally
used. When the body size and therefore body mass vary by several orders of
magnitude, one downside of abundance is that we mix individual that have
potentially really different meaning in terms of ecosystem properties such as
biomass. We therefore estimated the body mass of fish individuals according to
their body size. The body mass of individual fishes were then estimated with an
allometric law ($B_i = 0.01 \times l_i^{3.03}$, $B$ being the body mass in grams,
$l$ the body size in millimeters and $i$ the individual)
[@lleonart_removing_2000; @froese_cube_2006].

The species and community biomass were then computed for each sampling event.
As sampling effort, reported as the surface sampled during the sampling event,
varied across protocol type and sampling event, biomass was reported to the
surface sampled (in $m^{2}$). Community and species biomass were then expressed
in $gm^{-2}$.

Total biomass was characterized as the median total biomass over time. Temporal
biomass stability was defined as the inverse of biomass variability. Variability
was computed as Coefficient of Variation (CV) of the biomass ($CV_i = \sigma_i
\mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being respectively the standard
deviation and the temporal average of the biomass of the station $i$). Then, temporal
stability was equal to $CV^{-1} = \mu \sigma^{-1}$.

To characterize the mecanisms driving temporal stability, we decomposed
the community CV in synchrony and weighted average species CV
components according to @thibaut_understanding_2013:

@@

with $CV_{com}$ being the Coefficient of Variation of the community,
$\overline{CV_{sp}}$ the average coefficient of variation of population
pondered by biomass. $\phi$ is the synchrony, which is the ratio the total
community variance ($\sigma^2_{x_T}$) and the sum of population variance
($\sigma_{x_i}$) [@loreau_species_2008].

$$
\overline{CV_{sp}} = \sum_i{\frac{B_i {CV}_i}{\sum_i{B_i}}}
$$

@@

## Hozizontal & vertical diversity

### Horizontal diversity

####  Species richness

@@

### Vertical diversity

#### Trophic network inference

@@

Trophic interaction network was inferred from body size and ontogenic diet.
Body size is a reasonable approximation widely used to infer trophic
interactions [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan. For example, alvins feed on plankton while adults may feed on
algae or fishes. Accordingly,  each fish species was divided in nine body size
classes. Each body size class constituting a trophic species. A high congruence
has been found at trophic species scale between inferred interactions and
measured trophic interactions, but not at the species
scale [@jennings_weak_2001]. We then inferred trophic interactions between the trophic species (Figure
\@ref(fig:matmet) C).  We determined network structure in two steps: (1) build a
metaweb describing the trophic interactions between all trophic
species (Figure \@ref(fig:matmet) C), then (2) determine network of a given
community by subtracting the metaweb according to the trophic species present in
a given community (Figure \@ref(fig:matmet) D). We added to each network seven
resource nodes present in food diet database: detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos (grey nodes in Figure
\@ref(fig:matmet) C & D). The trophic network inference procedure is detailed in
supplementary informations.

#### Trophic network structure

In each network, we computed two network metrics that have been commonly used to
characterize network structure. Connectance which describes the number of
interactions compared to the number of possible links, computed as ($C = L /
N^2$, L and N being
resp. the number of links and of trophic species plus resource nodes). We then
computed the average trophic level, the average trophic level was weighted by
the biomass of each trophic species (excluding resource nodes for which we had
no biomass) ($\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$
and $B_i$ being resp. the trophic level and the biomass of the node $i$).

In resume, we infered binary trophic networks at the scale of trophic species
(i.e. body size classes by species) to take in account ontogenic diet shifts and
that fish-fish trophic interaction depends on body size ratio. On the other
side, the community metrics were defined at the species level, (i.e. species
richness, average species CV and synchrony).

@@

## Environment

@@

Habitat description, obtained from fish sampling data, describing the ecosystem
size namely the stream width, depth and position (altitude and
slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018; @tonkin_flow_2018; @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water and so, BOD is
also linked to physiological stress of fishes. As physico-chemical variables
were not recorded at the same place than fish communities, they were
interpolated with spatial stream networks [@isaak_applications_2014;
@hoef_ssn:_2014]. The interpolation procedure is detailed in supplementary
methods.

For each fish station, environmental variables were matched with the year of
fish sampling (except for water temperature for which the record began in 2006).
Then, the average temperature, flow, BOD, river width and depth were computed as
well as coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and
environmental variables are often collinear, a principal component analysis (PCA)
was performed on the scaled variables. The PCA explained `r tot_var_expl_pca`%
of the variance. An axis rotation was performed to maximize the representation
of the variables by the principal components (Fig. \@ref(fig:pca-rotated)).
Five principal components were used in the analysis, representing respectively
the (1) river size (river width & depth), (2) altitude, slope and decreasing
average temperature, (3) average enrichment (BOD) and CV of flow, (4) river size
CV (CV of river width and depth) and (5) CV of enrichment (CV of BOD). Further
informations about data transformation and filtering are detailed in
supplementary information.

@@

#### Summarise SEM statistical framework

Structural Equation Modeling are able to assess multivariate relationships,
directs and indirects between variables.  The metamodel describes how variables
are interrelated (Figure \@ref(fig:matmet) A). We built separated structural
equation model for total biomass and its temporal variability (Figure
\@ref(fig:sem-plot) A & B). Structural equation modeling was set up according to the
hypothesis described in introduction and detailed in supplementary information.
Structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary
informations). To account for heterogeneity of hydrographic basin, we set it as random effect on the intercept. Moreover, statistical model were checked for the presence of
multicollinearity (Supplementary information). The slopes reported in the main
text are standardized in order to compare their magnitude ($r_\delta = \beta
\frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope, $\sigma_x$
and $\sigma_y$ being respectively the standard deviation of the predictive
variable $x$ and of the response variable $y$). In the main text, only the
slopes which had $P$ value inferior to $0.05$ were reported (but see Table SXX).
Linear models and SEMs were computed with respectively the `nlme` and
`piecewieseSEM` R packages. The code used for the analysis are accessible on
github.

The structural equation models used for total biomass and its temporal stability
contains respectively 4 and 6 response variables, 35 and 26 direct paths.

#### Computation of indirect effects

The indirect effects of variables on total biomass and its temporal stability
was computed by multiplying slopes along the path. As an example, if we consider
$A \to B \to C$, the indirect effect of A on C is equal to $r_{AB} \times r_{BC}$.
Only significant direct slopes were entered in the calculation of
indirect effects. Finally total effects were computed as the sum of direct and
indirect effects.

@@
